---
title: "Video - R Regression"
output: html_notebook
---

Instructions:
Write the specified R  or provide the best response for each question.  I state it down in the exam, but I'll mention it here, too.  When testing the assumptions, DO NOT fix the problem and start all of the way back at the full model.  You will do that in your actual regression, but it may become very time consuming here.  That is not the point of this exam...  Understanding that you found an issue is the actual purpose of testing the assumptions.

When you remove variables, please place a _commment_ in the code, above the lm() or glm() stating which variable has been removed and why.

I wish you all well.


## 1. Read the RealEstateValue.xlsx file


```{r}
library(readxl)
RealEstateValue <- read_excel("/Users/beatrizmorales/Downloads/RealEstateValue.xlsx")
View(RealEstateValue)
```


## 2. Create a training dataset (70%) and a testing dataset (30%)
```{r}
library(caret)
set.seed(123)
training.samples <-RealEstateValue$PricePerSQFT |>
  createDataPartition(p=0.7, list=FALSE)
train_data <- RealEstateValue[training.samples,]
test_data  <- RealEstateValue[-training.samples,]  
test_data
```

## From this point, use the training dataset until told to use the test dataset.

## 3. Plot each independent variable (X-axis) against the dependent variable (Y-axis)
Note: The dependent variable is PricePerSQFT.  All values are continuous...no factors; thus, no dummy variables

```{r}
library(ggplot2)
library(rlang)
independent_vars <- setdiff(names(train_data), "PricePerSQFT")
for (var in independent_vars) {
  plot <- ggplot(train_data, aes_string(x = var, y = "PricePerSQFT")) +
    geom_point(alpha = 0.6) +
    geom_smooth(method = "lm", se = FALSE, color = "red") +
    labs(title = paste("PricePerSQFT vs", var),
         x = var,
         y = "PricePerSQFT") +
    theme_minimal()
  print(plot)
}
```

## 4. Do any variables seem to have a 2nd (parabola - one bend) or 3rd degree (2 bends) polynomial relationship with the DV?
----
The independent variable (HouseAge) seem to have a possible 2nd degree relationship, as it shows a curved pattern with the dependent variable PricePerSQFT. The midaged houses have higher prices than the very old or new houses. The other independent variables appear to have no indication of a second or third degree polynomial trend.

----

## 5. Create a correlation matrix using all of the predictors.

```{r}
predictors <- train_data[, setdiff(names(train_data), "PricePerSQFT")]
cor_matrix <- cor(predictors)
print(cor_matrix)
```

## 6. Which, if any, of the predictors (IVs) are correlated above 0.4 or below -0.4 with other predictors?  What is the correlation coefficient between the two variables?
Note: Don't do anything about it, yet.  Just observe and document them below.
----
A few of the predictors show moderate to strong correlations with each other. The ToTransitStation and Longitude has the strongest correlation coefficient of -0.78856658. Next the ToTransitStation and NumConvenienceStores with a -0.58667035. The next two correlations decrease, NumConvenienceStores and with both Latitude 0.41937699 and Longitude 0.41755719.
----


## 7. Using either lm() or glm() for linear regression, regress PricePerSQFT on all other variables (full model) and any polynomial representations of the variables, depending on your answer to #4.

Display the results of the regression.

```{r}
train_data$HouseAge2 <- train_data$HouseAge^2
full_model <- lm(PricePerSQFT ~ HouseAge + HouseAge2 + ToTransitStation + 
                   NumConvenienceStores + Latitude + Longitude, 
                 data = train_data)
summary(full_model)
```

## 8. Calculate RMSE for the full model
```{r}
residuals_full <- residuals(full_model)
rmse <- sqrt(mean(residuals_full^2))
print(rmse)
```

## 9. Use backward elimination to remove one variable at a time, if needed.  After each regression with an eliminated variable:
  a. calculate RMSE for the model,
  b. use anova() to determine if each reduced model fits the data as well as the full model () and 
  c. calculate the AIC() comparing the reduced model(s) to the full model for each successive reduced model.

Note: Create as many separate chunks as you need and keep removing variables.  Number (or code in some way) each reduced model so you will have a fit variable for the full model and each reduced model.

STOP WHEN ALL PREDICTORS ARE STATISTICALLY SIGNIFICANT


#### Regress on predictors after eliminating a variable one at a time and show results
```{r}
model_0 <- full_model
model_1 <- lm(PricePerSQFT ~ HouseAge + HouseAge2 + ToTransitStation +
                NumConvenienceStores + Latitude, data = train_data)
rmse_1 <- sqrt(mean(residuals(model_1)^2))
print(rmse_1)
anova(model_1, model_0)
aic_1 <- AIC(model_1)
print(aic_1)
summary(model_1)
```


#### Calculate RMSE
```{r}
residuals_final <- residuals(model_1)
rmse_final <- sqrt(mean(residuals_final^2))
print(rmse_final)
```

#### compare Fit using anova() - full versus most recent fit variable
```{r}
anova(model_1, model_0)
```

#### compare fit using AIC()- full versus most recent fit variable
```{r}
aic_full <- AIC(model_0)
aic_reduced <- AIC(model_1)
print(paste("Full model AIC:", aic_full))
print(paste("Reduced model AIC:", aic_reduced))
```


## 10. Create a table with each fit on a separate row.  Columns should be Model Name, RMSE, anova(), AIC()
-----
Model Name      Adj. R-square     RMSE          anova()         AIC()
full.model       0.5863          76.91         ----            ~3371.717
reduced.model    0.5877          75.99        p=0.8104(NS)      3371.717
. . .

----

## 11. Which model do you choose and why?
----
The model I would choose is the reduced.model, due to its predictors being statistically significant at (p<0.05), compared to the full.model at (p=0.81). The RMSE is lower at 75.99 on the reduced.model than the full.model at 76.91. The model with the lowest RMSE is always the best model to go with.
----

## Note: When testing the assumptions, multicollinearity, and outliers, for this exam, DO NOT take action and start over with the regression...  Just remember that in 'real life,' you should go back to the full model and start again once you remediate each violation.  That is why on most of the related questions, I ask you what you should do.  DON'T DO IT or this exam could take forever to complete.

## 12. For the chosen model, run the plot(...) and comment on all graphs that were discussed in the videos

```{r}
reduced.model <- model_1
plot(reduced.model)
```

## 13. For the chosen model, assess the normality of the residuals
```{r}
qqnorm(residuals(reduced.model))
qqline(residuals(reduced.model), col = "red")
shapiro.test(residuals(reduced.model))
```

## 14. Does there seem to be a violation of the normality assumption?  Don't fix it in this exam, just comment
----
There seems to be some violation of the normality assumption based on the Q-Q plot. The residuals are aligned with the diagonal line, but on both ends there seems to be some deviation indicating mild skewness.
----

##15. For the chosen model, test for independence of errors.  
```{r}
library(lmtest)
dwtest(reduced.model)
acf(residuals(reduced.model), main = "ACF Plot of Residuals")
```

## 16. Does there seem to be a violation of independence?  What other way should the assumption of independence be assessed?
----
It does not seem to be a violation of independence since the residuals are not autocorrelated and the ACF plot confirms no spike falling within the blue dash line.
----

## 17. For the chosen model, test for linearity?
```{r}
plot(reduced.model$fitted.values, residuals(reduced.model),
     xlab = "Fitted Values", ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, col = "red")
library(car)
crPlots(reduced.model)
```

## 18.  Does what do you see in the results that indicate a violation or not a violation of the assumption?
----
The results indicate there is not a violation of the assumption for the reduced model. The plot show the residuals to be scattered around the horizontal line without a clear pattern.
----


## 19. For the chosen model, test the assumption of homoscedasticity.
```{r}
ncvTest(reduced.model)
spreadLevelPlot(reduced.model)
```

## 20. Does there seem to be a violation of the assumption of homoscedasticity? What evidence do you see?
----
There is evidence of a violation of the assumption of homoscedasticity, since the p value is less that 0.05. With a p=0.003 and a Chisquare value of 8.6476, it reflects how the residuals are not constant.
----

## 21. For the chosen model, test for multicollinearity.
```{r}
vif(reduced.model)
```

## 22. Do you see any evidence of multicollinearity?  Don't do it, but what should you do to fix it?
----
There is evidence present in reduced model test for multicollinearity. The HouseAge and HouseAge2 are both above 10 which exceed the threshold.
----

## 23. For the chosen model, test for outliers?
```{r}
outlierTest(reduced.model)
```

## 24. Do you see any evidence of outliers?  If so and it is bad data, what should you do to eliminate its impact?
----
There is evidence of outliers by looking at the Bonferroni p value for 193. If it is confirmed to be bad data or data entry error, it could be removed or corrected. By correcting the bad data it should help minimize the impact.
----

#25. Run an All Subsets Regression using Adjusted R-square as your measure.
```{r}
library(leaps)
subset_model <- regsubsets(PricePerSQFT ~ HouseAge + HouseAge2 + ToTransitStation + 
                             NumConvenienceStores + Latitude, 
                           data = train_data, nbest = 1)
subset_summary <- summary(subset_model)
subset_summary$adjr2
```

## 26. Which model write the formula you should use based on the all subsets model (e.g. Y ~ X1 + X2 + X3), except with the actual variable names.
----
(PricePerSQFT ~ HouseAge + HouseAge2 + ToTransitStation + Latitude)

----

## 27. Using the testing dataset, run the chosen model (formula for the chosen model), display the summary() of the fit, and calculate the RMSE comparing the DV values in the testing dataset with the fit of the model run on the testing dataset.
```{r}
test_data$HouseAge2 <- test_data$HouseAge^2
test_predictions <- predict(reduced.model, newdata = test_data)
rmse_test <- sqrt(mean((test_data$PricePerSQFT - test_predictions)^2))
print(rmse_test)
summary(reduced.model)
```

